---
title: "Participation & global conflict"
format: html
editor: visual
---

```{R}
install.packages("tidytuesdayR")

tuesdata <- tidytuesdayR::tt_load('2021-07-27')
tuesdata <- tidytuesdayR::tt_load(2021, week = 31)

olympics <- tuesdata$olympics

olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')
```

## Data wrangling analysis 1 - AC

```{r}
#So the winter and summer were held in the same year up until 1992. 
#In editing phase I'd like to
#We are going to look at the data from 1896 - 1992. first we filter
library(dplyr)
library(ggplot2)
library(stringr)
subolympics <- olympics %>% filter(year < 1994)
# create data set that sums the countries by year. 
countries <- subolympics %>%
  group_by(noc,year,sex) %>%
  summarise(participantnum = n(), .groups = "drop")

```

## Data wrangling analysis 2 - AC

```{r}
## Analysis 2 - AC: Shiny app: Input your age, weight, height and gender. Spits out who medaled in what sport with those characteristics.
#essentially we need to omit anyone with out the data we need. 
medalist <- olympics %>% na.omit()
```

#Analysis 1 - AC: How does global conflict influence the number of nations who participate in the Olympics?

```{r fig.align="center", echo=FALSE, fig.width=14}
#split the data into two frames summer and winter
conflict = "Start of Iran and Iraq War, 
                          Soviet-Afgan War, 
                          Cold War tension"

ggplot(countries, aes(x=year, y=participantnum, fill = sex)) +
  geom_col() +
  scale_x_continuous(n.breaks = 49) +
  scale_y_continuous(n.breaks = 35, expand = c(0,0)) +
  theme(axis.text.x = element_text(angle=90, vjust = 0.5, hjust = 1))+
  theme(legend.position = "bottom") +
  labs(title = "Influence of global conflict on the number of olympic participants", x="Year of olympic games", y= "Total number of participants")+
  theme(plot.title = element_text(size = 16, face= "bold", hjust = 0.5)) +
     geom_label(label = "World War 1", x=1916, y = 4700, 
             label.padding = unit(0.55, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "#69b3a2")+
      geom_segment(aes(x = 1916, y = 4350, xend = 1916, yend = 500),
                  arrow = arrow(length = unit(0.5, "cm")))+

     geom_label(label = "Great Depression", x=1932, y = 8000, 
             label.padding = unit(0.55, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "#69b3a2") +
    geom_segment(aes(x = 1932, y = 7650, xend = 1932, yend = 3500),
                  arrow = arrow(length = unit(0.5, "cm")))+
  
  geom_label(label = "World War 2", x=1942, y = 8000, 
             label.padding = unit(0.55, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "#69b3a2")+
     geom_segment(aes(x = 1942, y = 7650, xend = 1942, yend = 0),
                  arrow = arrow(length = unit(0.5, "cm")))+
  
     geom_label(label = "Suez Conflict", x=1956, y = 10000, 
             label.padding = unit(0.55, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "#69b3a2")+
    geom_segment(aes(x = 1956, y = 9650, xend = 1956, yend = 6500),
                  arrow = arrow(length = unit(0.5, "cm")))+
  

     geom_label(label = str_wrap(conflict,20), x=1980, y = 13000, 
             label.padding = unit(0.55, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "#69b3a2") +
   geom_segment(aes(x = 1980, y = 12000, xend = 1980, yend = 9500),
                  arrow = arrow(length = unit(0.5, "cm")))
  


```

## Analysis 2 - AC: How do your physical characteristics match up with historical olympians?

```{r setup, include=FALSE}
library(flexdashboard)
```

---
title: "How do your physical characteristics match up with historical olympians"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

Column {data-width=1000}

```{r picture, echo=F, outwidth='100%'}

library(dplyr)
library(tibble)
library(shiny)
library(purrr)
library(DT)
library(shinydashboard)
library(rsconnect)
library(ggplot2)
library(shinythemes)
library(shinyWidgets)
library(shinydashboard)
library(shinydashboardPlus)



# start of application
sidebar <-  dashboardSidebar(
      width = 250,
      sidebarMenu(id = "tabs", 
       menuItem("Dashboard",
                tabName = "dashboard", 
                selected = TRUE, 
                startExpanded = TRUE,
       menuItem("About us", tabName = "about"),
       menuItem("Olympian's physical characteristics", tabName = "subitem1"),
       menuItem(HTML("Help desk:<br/> imperial to metric conversion"),  tabName = "Imperial")
       
  )
)
)
  
 body <- dashboardBody(
  tabItems(
  tabItem(tabName = "about",
          userBox(
          title = userDescription(
          title = "Addison Carroll",
          subtitle = "PhD Animal Science",
          type = 2,
          image = "https://lh3.googleusercontent.com/drive-viewer/AK7aPaBIIZV2XU74yuhj4R5lPX64CaCUVSttH8XY3ZrIJ0fqAhA5U7rre_nbHLA_xB_89CGqvb7LTTf_w71QerBAMYIinqaJWw=s1600?source=screenshot.guru",
        ),
        status = "primary",
        gradient = TRUE,
        background = "black",
        collapsible = FALSE,
        width = 6,
        boxToolSize = "xl",
        "Hi! I'm Addison a PhD with a focus in dairy cattle nutrition. I enjoy metalcore music,  my Aussie Bear, and exploring data for hidden relationships.",
        footer = NULL
        ),
         userBox(
          title = userDescription(
          title = "Sydney Graham",
          subtitle = "PhD Plant Breeding and Genetics",
          type = 2,
          image = "ACwJerseys.png",
        ),
        status = "primary",
        gradient = TRUE,
        background = "maroon",
        collapsible = FALSE,
        width = 6,
        boxToolSize = "xl",
        "Some text here!",
        footer = NULL
      )),
 tabItem(tabName = "subitem1", 
    fluidPage(
      titlePanel(div("Influence of physical characteristics on medaling in Olympic games.. See how you stack up!",
                 style = "color: #000000")),
       fluidRow(
            column(12,
        chooseSliderSkin("Shiny", color = "red"),
          sliderInput("age",
                    "Age range", 
                    min= min(medalist$age),
                    max= max(medalist$age), 
                    value = c(min(medalist$age), max(medalist$age)),
                    width = '100%'),
                    br()
    ),
         
        column(6,
        wellPanel(style = "background-color: #960018; color: #FFFFFF;",
        selectInput("sex",
                    "Gender:",
                    c("All",
                      unique(as.character(medalist$sex)))))
        ),
         column(6,
          wellPanel(style = "background-color: #960018; color: #FFFFFF;",
          selectInput("season",
                    "Summer or Winter season:",
                    c("All",
                      unique(as.character(medalist$season))))),
                       br()
      ),

      column(6,
        wellPanel(style = "background-color: #960018; color: #FFFFFF;",
        pickerInput(inputId = "weight",
                    label ="Weight (kg)",
                    choices = c("All", unique(sort(medalist$weight)),
                    multiple = FALSE
        )))

     ),
      column(6,
             wellPanel(style = "background-color: #960018; color: #FFFFFF;",
             pickerInput(inputId = "height",
                    label ="Height (cm)",
                    choices = c("All", unique(sort(medalist$height))),
                    multiple = FALSE
         )
        )
    ),
    tags$style('.container-fluid { background-color: #FFFFFF;}'),
     tabsetPanel(tabPanel("outcome", DT::dataTableOutput("table")))
       )
    )
       ),
    
  tabItem(tabName = "Imperial",
        h2("All data is in metric.. Here is the conversion for us dummies who don't use it!"),
            wellPanel(style = "background-color: grey;",
                      numericInput("myInput1", label = "Type in height in inches", value = 1),
                       uiOutput("ui1")),
             wellPanel(style ="background-color: #960018;",
                       numericInput("myInput2", label = "Type in weight in lbs", value = 1),
                       uiOutput("ui2"))
)
)
)


ui <- dashboardPage(skin = "midnight", 
  dashboardHeader(title =  span(tagList(icon("earth-americas"), "olympics"))),
  sidebar,
  body
)


server <- function(input, output, session) {

  # Filter data based on selections
  output$table <- DT::renderDataTable(DT::datatable({
    data <- medalist[medalist$age >= input$age[1] & medalist$age <= input$age[2],]
    if (input$sex != "All") {
      data <- data[data$sex == input$sex,]
    }
    if (input$season != "All") {
      data <- data[data$season == input$season,]
    }
    if (input$weight != "All") {
      data <- data[data$weight == input$weight,]
    }
    if (input$height != "All") {
      data <- data[data$height == input$height,]
    }
    data 
  }) %>%
      DT::formatStyle(colnames(data), target = 'cell', backgroundColor = 'white', opacity = 1) 
  
  ) 
 
  
  output$ui1 <- renderUI( {
    tagList(
      numericInput("obs1", "Height in cm", value = input$myInput1*2.54))
  })
    output$ui2 <- renderUI( {
    tagList(
      numericInput("obs1", "Weight in kg", value = input$myInput2*0.453592))
    })
    observe({print(input$tabs)})
}


# Run the application
shinyApp(ui, server)
```

#Conclusions
